\documentclass[aspectratio=169,11pt]{beamer}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{xcolor}
\usepackage{ragged2e}
\usepackage{colortbl}
\usepackage{array}
\usepackage{hyperref}
\usepackage{adjustbox}

\usetikzlibrary{shapes.geometric, arrows.meta, positioning, calc, backgrounds, decorations.pathreplacing, shadows, fadings}
\pgfplotsset{compat=1.18}

% --- Color Palette: Warm Professional ---
\definecolor{DeepNavy}{HTML}{2E4057}
\definecolor{Teal}{HTML}{048A81}
\definecolor{WarmOrange}{HTML}{E85D04}
\definecolor{SoftPurple}{HTML}{9D4EDD}
\definecolor{WarmGray}{HTML}{6C757D}
\definecolor{LightGray}{HTML}{E9ECEF}
\definecolor{Cream}{HTML}{FBF8F1}
\definecolor{DeepRed}{HTML}{D62828}
\definecolor{Gold}{HTML}{D4A03A}
\definecolor{SoftWhite}{HTML}{FAFAFA}

% Beamer colors
\setbeamercolor{normal text}{fg=DeepNavy, bg=SoftWhite}
\setbeamercolor{structure}{fg=DeepNavy}
\setbeamercolor{alerted text}{fg=DeepRed}
\setbeamercolor{example text}{fg=Teal}
\setbeamercolor{frametitle}{fg=DeepNavy, bg=SoftWhite}
\setbeamercolor{title}{fg=DeepNavy}
\setbeamercolor{subtitle}{fg=WarmGray}
\setbeamercolor{block title}{bg=DeepNavy, fg=SoftWhite}
\setbeamercolor{block body}{bg=LightGray, fg=DeepNavy}
\setbeamercolor{itemize item}{fg=WarmOrange}
\setbeamercolor{itemize subitem}{fg=Gold}

% Fonts
\usefonttheme{professionalfonts}
\setbeamerfont{title}{size=\huge, series=\bfseries}
\setbeamerfont{frametitle}{size=\Large, series=\bfseries}

% Strip chrome
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{headline}{}

% Minimal footline
\setbeamertemplate{footline}{%
    \hfill
    \begin{beamercolorbox}[wd=3cm,ht=2.5ex,dp=1ex,right,rightskip=0.6cm]{page number}
        \usebeamercolor[fg]{WarmGray}\scriptsize\insertframenumber
    \end{beamercolorbox}
    \vspace{0.3cm}
}

% Bullets
\setbeamertemplate{itemize item}{\footnotesize\raisebox{0.3ex}{\tikz\fill[WarmOrange] (0,0) circle (0.45ex);}}
\setbeamertemplate{itemize subitem}{\footnotesize\raisebox{0.3ex}{\tikz\fill[Gold] (0,0) circle (0.35ex);}}

% Custom commands
\newcommand{\transitionslide}[2]{%
    {\setbeamercolor{normal text}{fg=SoftWhite,bg=DeepNavy}
    \begin{frame}[plain]\vfill\begin{center}
        {\huge\bfseries\textcolor{Gold}{#1}}\\[0.6cm]
        {\large\textcolor{LightGray}{#2}}
    \end{center}\vfill\end{frame}}
}

\begin{document}

\transitionslide{Application}{GIV in the Israeli Labor Market}

\begin{frame}
    \frametitle{The Israeli Tech Sector is highly granular}

    \vspace{0.5cm}
    {\Large
        The Israeli high-tech industry is dominated by a few massive multinational R\&D centers.
    }

    \vspace{1cm}
    \begin{itemize}
        \item \textbf{High Concentration:} Companies like Intel, Google, Microsoft, and NVIDIA account for a substantial, disproportionate share of tech employment.
        \item \textbf{A Heavy Tail:} The size distribution of tech employers in Israel exhibits a strong power law, yielding a high \textbf{excess Herfindahl index ($h$)}.
        \item \textbf{Macro vs. Micro:} It is difficult to separate general Israeli macroeconomic conditions (e.g., security shocks, interest rates) from the hiring dynamics of these specific giants.
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Applying GIV to Israeli tech wages}

    \vspace{0.5cm}
    {\Large
        How do we isolate the causal spillover of these tech giants on aggregate Israeli wages?
    }

    \vspace{1cm}
    \begin{itemize}
        \item \textbf{The Idiosyncratic Shocks ($u_{it}$):} We exploit global shocks specific to the multinationals that are arguably exogenous to the Israeli macroeconomy (e.g., global restructuring at Intel, a worldwide AI hiring boom at NVIDIA).
        \item \textbf{The Instrument ($z_t$):} Construct the GIV by taking the size-weighted employment shocks of these giants and projecting out the common Israeli macroeconomic factor loadings.
        \item \textbf{The 2SLS Estimation:} Regress aggregate Israeli wages on the component of labor demand driven \emph{only} by these isolated, granular tech shocks.
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Takeaways: GIV unconfounds local macro dynamics}

    \vspace{1cm}
    {\Large
        The GIV method solves the simultaneity problem in concentrated labor markets.
    }

    \vspace{1cm}
    \begin{itemize}
        \item \textbf{Causal Wage Spillovers:} We can measure exactly how much a hiring freeze at a major tech firm depresses aggregate local wages, separated from a general economic downturn.
        \item \textbf{Beyond the US:} The method is perfectly suited for small, open economies like Israel, where granular actors often drive aggregate volatility.
        \item \textbf{The Broad Lesson:} We don't need perfect natural experiments; the everyday idiosyncratic variance of the giants provides all the identifying variation we need.
    \end{itemize}
\end{frame}

\end{document}